---
title: "PM566 Final Project"
author: "YINA LIU"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>

This is my PM566 Final Project website. 
<br>

```{r setup1, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```

<br>

## First step

First I will source any necessary code

```{r load-data}
source("process_COVID_data.R")
```

<br>

## Second step

Then I will add some code to create the **plotly** figures


```{r plot1, class.source="code-r-small"}

p1_scatter <- cv_states_today %>% 
  plot_ly(x = ~pop_density, y = ~deathsper100k,
          type = 'scatter', mode = 'markers', color = ~state,
          size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5),
          hoverinfo = 'text',
          text = ~paste( paste(state, ":", sep=""), paste(" Cases per 100k: ", per100k, sep="") , paste(" Deaths per 100k: ",
                        deathsper100k, sep=""), sep = "<br>")) %>%
  layout(title = "Population-normalized COVID-19 deaths vs. population density",
                  yaxis = list(title = "Deaths per 100k"), xaxis = list(title = "Population Density"),
         hovermode = "compare")

# filter out "District of Columbia"
cv_states_today_scatter <- cv_states_today %>% filter(state!="District of Columbia")

p2_scatter <- cv_states_today_scatter %>% 
  plot_ly(x = ~pop_density, y = ~deathsper100k,
          type = 'scatter', mode = 'markers', color = ~state,
          size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5),
          hoverinfo = 'text',
          text = ~paste( paste(state, ":", sep=""), paste(" Cases per 100k: ", per100k, sep="") , paste(" Deaths per 100k: ",
                        deathsper100k, sep=""), sep = "<br>")) %>%
  layout(title = "Population-normalized COVID-19 deaths vs. population density",
                  yaxis = list(title = "Deaths per 100k"), xaxis = list(title = "Population Density"),
         hovermode = "compare")
```

<br>

## Third step {.tabset}

Create tabs to display each figure

### Scatterplot: with DC

```{r p1}
p1_scatter
```

### Scatterplot: without DC

```{r p2}
p2_scatter
```


## Assignment 5 {.tabset}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, include = FALSE}
library(data.table)
library(dtplyr)
library(dplyr)
library(lubridate)
library(leaflet)
library(ggplot2)
library(knitr)
library(kableExtra)
library(lubridate)
library(qwraps2)
library(plotly)
```
```{r, include = FALSE}
# read in the datasets
all_data <- data.table::fread("/Users/apple1/Desktop/USC\ Biostatistics/PM566/HW/PM566\ Midterm\ Project/data/Crime_Data_from_2020_to_Present.csv")
bul <- all_data[which(all_data$'Crm Cd Desc' == 'BURGLARY')]
```


```{r, include = FALSE}
# check the dimensions, headers and footers
dim(bul)
head(bul)
tail(bul)

# check the variable names and variable types
str(bul)

#Assign and combine desenct name to desenct code
bul$`Vict Descent`[bul$`Vict Descent`=='A'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='B'] <- 'Black'
bul$`Vict Descent`[bul$`Vict Descent`=='C'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='D'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='F'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='G'] <- 'Guamanian'
bul$`Vict Descent`[bul$`Vict Descent`=='H'] <- 'Hispanic/Latin'
bul$`Vict Descent`[bul$`Vict Descent`=='I'] <- 'AIAN'
bul$`Vict Descent`[bul$`Vict Descent`=='J'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='K'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='L'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='O'] <- 'Other'
bul$`Vict Descent`[bul$`Vict Descent`=='P'] <- 'Pacific Islander'
bul$`Vict Descent`[bul$`Vict Descent`=='S'] <- 'Samoan'
bul$`Vict Descent`[bul$`Vict Descent`=='U'] <- 'Hawaiian'
bul$`Vict Descent`[bul$`Vict Descent`=='V'] <- 'Asian'
bul$`Vict Descent`[bul$`Vict Descent`=='W'] <- 'White'
bul$`Vict Descent`[bul$`Vict Descent`=='Z'] <- 'Asian'

# transform character to date and extract month from occurance date
bul$`DATE OCC` <- as.Date(substr(bul$`DATE OCC`,1,10),"%m/%d/%Y")
bul$month_occurance <- month(bul$`DATE OCC`)

# take a closer look at the key variables. 

# check the `Vict Age`, `Vict Sex` and `Vict Descent` variables.
summary(bul$`Vict Age`)
table(bul$`Vict Sex`)
table(bul$`Vict Descent`)

# check how many Geographic Areas in LA has been evaluated 
length(unique(bul$AREA))
length(unique(bul$`AREA NAME`))

# check the range of lattitude and longitude.
summary(bul$LAT)
summary(bul$LON)

```

```{r, include = FALSE}
# check for missing values
age_mv <- sum(bul$`Vict Sex`=='') + sum(bul$`Vict Sex`=='X')
age_mv/length(bul$`Vict Sex`)

des_mv <- sum(bul$`Vict Descent`=='') + sum(bul$`Vict Descent`=='X')
des_mv/length(bul$`Vict Descent`)

# check for impausible values and its proportion
sum(bul$`Vict Age`== 0)
sum(bul$`Vict Age`== 0)/length(bul$`Vict Age`)

sum(bul$LAT== 0)
sum(bul$LAT== 0)/length(bul$LAT)
sum(bul$LON== 0)/length(bul$LON)

# remove the missing and impausible values
bul <- bul[which(bul$`Vict Sex`!='')]
bul <- bul[which(bul$`Vict Sex`!='X')]
bul <- bul[which(bul$`Vict Descent`!='')]
bul <- bul[which(bul$`Vict Descent`!='X')]

bul <- bul[which(bul$`Vict Age`!=0)]
summary(bul$`Vict Age`)

bul <- bul[which(bul$LAT!=0)]
bul <- bul[which(bul$LON!=0)]
summary(bul$LAT)
summary(bul$LON)
```

### Boxplot for age
```{r, echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p<-ggplot(bul,aes(y=`Vict Age`))+geom_boxplot()
ggplotly(p)
```

### Barchart for sex
```{r,  echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p <- ggplot(bul,aes(x =`Vict Sex`)) + geom_bar(stat = 'count', fill = 'lightblue', colour = 'darkblue') + labs(title = "Numbers by sex")
ggplotly(p)
```

### Barchart for descent
```{r,  echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p<-ggplot(bul,aes(x =`Vict Descent`)) + geom_bar(stat = 'count', fill = 'lightpink', colour = 'darkred') + labs(title = "Numbers by descent")
ggplotly(p)
```

### Boxplot for age by sex
```{r,  echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p<-ggplot(bul,aes(x=factor(`Vict Sex`),y=`Vict Age`))+geom_boxplot() 
ggplotly(p)
```

### Boxplot for age by descent
```{r, echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p<-ggplot(bul,aes(x=factor(`Vict Descent`),y=`Vict Age`))+geom_boxplot() 
ggplotly(p)
```

### Barchart for sex by descent
```{r, echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p<-ggplot(bul,mapping = aes(x = `Vict Descent`, fill = `Vict Sex`))+
  geom_bar()+
  labs(title = "Vict Descent Category by Sex")+
  scale_fill_brewer(palette = "Set3")
ggplotly(p)
```

### Barchart for AREA
```{r, echo = FALSE, out.width = "588", out.height = "378",warning=FALSE}
p<-ggplot(bul,aes(x =AREA)) + geom_bar(stat = 'count', fill = 'lightyellow', colour = 'orange') + labs(title = "Numbers by Area")
ggplotly(p)
```

### Leaflet for gender
```{r, echo = FALSE,warning=FALSE}
colorplot <- if_else(bul$`Vict Sex`=="F","red","blue")
leaflet(bul) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~LAT, lng=~LON, radius = 0.01, color = colorplot, fillColor = colorplot ) 
```

### Line plot for occurace date except for October -- total
```{r, echo = FALSE,warning=FALSE}

bul1 <- bul[which(bul$month_occurance!=10)]

p<-ggplot(bul1,aes(x =month_occurance)) + geom_line(stat = 'count') +geom_point(stat = 'count')+scale_x_continuous(breaks=seq(0, 9, 1)) 
ggplotly(p)
```
### Line plot for occurace date except for October -- by gender
```{r,echo = FALSE,warning=FALSE}
p<-ggplot(bul1,aes(x =month_occurance, color = `Vict Sex`)) + geom_line(stat = 'count') +geom_point(stat = 'count')+scale_x_continuous(breaks=seq(0, 9, 1)) 
ggplotly(p)
```

### Line plot for occurace date except for October -- by descent
```{r,echo = FALSE,warning=FALSE}
p<-ggplot(bul1,aes(x =month_occurance, group=`Vict Descent`, color = `Vict Descent`)) + geom_line(stat = 'count') +geom_point(stat = 'count')+scale_x_continuous(breaks=seq(0, 9, 1)) 
ggplotly(p)
```

### Line plot for occurace date except for October -- by area
```{r,echo = FALSE,warning=FALSE}
p <- ggplot(bul1,aes(x =month_occurance, group=`AREA NAME`, color = `AREA NAME`)) + geom_line(stat = 'count') +geom_point(stat = 'count')+scale_x_continuous(breaks=seq(0, 9, 1)) 

ggplotly(p)
```

### Summary table for age, sex and descent
```{r, results = "asis", echo = FALSE, warning=FALSE}

options(qwraps2_markup = "markdown")

summary1 <-
  list("Victim Age" =
       list("min"       = ~ min(`Vict Age`),
            "max"       = ~ max(`Vict Age`),
            "mean (sd)" = ~ qwraps2::mean_sd(`Vict Age`)),
       
       "Victim Gender" =
       list("Male" = ~ qwraps2::n_perc0(`Vict Sex` == 'M'),
            "Female"  = ~ qwraps2::n_perc0(`Vict Sex` == 'F'))
       )
whole <- summary_table(bul, summary1, by = c("Vict Descent"))
whole
```

### Summary table for area and time
```{r, results = "asis", echo = FALSE, warning=FALSE}
bul$area_all <- paste0(bul$AREA,': ',bul$`AREA NAME`)

area_t <- xtabs(~area_all + month_occurance,data=bul)
area <- as.data.frame(cbind(area_t, total = margin.table(area_t,1)))
setDT(area, keep.rownames = TRUE)

library(DT)
areatb <- area %>%  select(rn,`1`,`2`,`3`,`4`,`5`,`6`,`7`,`8`,`9`,`10`, total)
datatable(areatb)


#kable(area, "html") %>%
#  kable_styling(bootstrap_options = c("striped", "hover"))

```